
---
title: "kbtbr"
output: github_document
---

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![R-CMD-check](https://github.com/CorrelAid/kbtbr/workflows/R-CMD-check/badge.svg)](https://github.com/CorrelAid/kbtbr/actions)
[![Codecov test coverage](https://codecov.io/gh/CorrelAid/kbtbr/branch/main/graph/badge.svg)](https://codecov.io/gh/CorrelAid/kbtbr?branch=main)
<!-- badges: end -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

`kbtbr` is a wrapper for the [KoBoToolbox APIs](https://support.kobotoolbox.org/api.html). It focuses on API v2 but also makes use of v1 if required (currently no dependencies on v1). `kbtbr` not only allows you to pull answers to your surveys directly into your R session but also lets you create and clone assets and import XLS forms straight from your R console. Finally, it provides flexible low-level functions to implement functionalities currently missing from the package yourself. 

# Installation

`kbtbr` is not on CRAN yet. You can install the current version from GitHub: 

```{r installation, eval=FALSE}
remotes::install_github("CorrelAid/kbtbr")
```

Install the development version (unstable!)

```{r installation-dev, eval=FALSE}
remotes::install_github("CorrelAid/kbtbr", ref = "dev")
```

# Get started

```{r eval=FALSE}
library(kbtbr)
# replace with https://kobo.humanitarianresponse.info for the humanitarian server or your own if you self-host
base_url_v2 <- "https://kf.kobotoolbox.org" 
token <- Sys.getenv("KBTBR_TOKEN")

kobo <- Kobo$new(base_url_v2, base_url_v1, token)
kobo$get_surveys()
```

See the documentation for more! 

# Documentation

Documentation is available as a [`pkgdown`](https://pkgdown.r-lib.org/) website:

- [stable version](https://correlaid.github.io/kbtbr/) 
- [development version](https://correlaid.github.io/kbtbr/dev/)

# Contributing to kbtbr

Please refer to the [CONTRIBUTING.md](https://github.com/correlaid/kbtbr/blob/main/.github/CONTRIBUTING.md).
