---
title: "Get started"
output:  rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started}
  %\VignetteEngine{knitr::knitr}
  %\VignetteEncoding{UTF-8}
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)

# properly set up vcr
library(vcr)
library(dplyr)
vcr_dir <- "./assets"
if (!nzchar(Sys.getenv("KBTBR_TOKEN"))) {
    if (dir.exists(vcr_dir)) {
        # Fake API token to fool our package
        Sys.setenv("KBTBR_TOKEN" = "fakebearertoken")
    } else {
        # If there's no mock files nor API token, impossible to run tests
        stop("No API key nor cassettes, tests cannot be run.",
            call. = FALSE
        )
    }
}

invisible(vcr::vcr_configure(
    dir = vcr_dir,
    filter_request_headers = list(Authorization = "fakebearertoken")
))
vcr::check_cassette_names()
```

# Overview
kbtbr is a wrapper for the [KoBoToolbox APIs](https://support.kobotoolbox.org/api.html). It focuses on API v2 but also makes use of v1 if required.

kbtbr uses [R6](https://r6.r-lib.org/articles/Introduction.html) classes. This means it has an object-oriented user interface which might feel unfamiliar to R users at first. But don't worry - you should get used to it quickly!



# Setup
## Authentication

Get an API token using any of the methods described in the [KoBoToolbox documentation](https://support.kobotoolbox.org/api.html#getting-your-api-token).
Then store it as an environment variable by opening and editing your `.Renviron`:


```{r eval=FALSE}
usethis::edit_r_environ()
```

Add your token on a new line as follows:

```
KBTBR_TOKEN=yourapitokenhere
```
Save the file and restart R/RStudio.


Of course, you can also use local environment files or any other method to securely store the API token. 

## Base URL
In order to use kbtbr, you need to know the _base URLs_ of your KoBoToolbox server (cf. [here](https://support.kobotoolbox.org/api.html#using-the-api)). There are two _base URLs_, one for API version 2 and one for version 1. 
To use the kbtbr package, you only need to specify the _base URL_ for version 2. However, certain functionalities that rely on the version 1 API will require its _base URL_. 

If you are using one of the [two hosted instances](https://support.kobotoolbox.org/server.html), those are the base URLs:


```{r echo=FALSE, eval=TRUE}
baseurls <- tibble::tribble(
  ~version, ~`Humanitarian Server (OCHA)`, ~`Non-Humanitarian Server`,
  "v2", "https://kobo.humanitarianresponse.info", "https://kf.kobotoolbox.org",
  "v1", "https://kc.humanitarianresponse.info", "https://kc.kobotoolbox.org" 
)

knitr::kable(baseurls)

```

## Creating the Kobo instance 

With the right base URL(s), we can create an instance of the `Kobo` class. The `base_url_v1` argument is optional.

```{r kobo}
library(kbtbr)
library(dplyr)
base_url_v2 <- "https://kobo.correlaid.org" # replace with your base URL for v2
base_url_v1  <- "https://kc.correlaid.org" # replace with your base URL for v1 
token <- Sys.getenv("KBTBR_TOKEN")

kobo <- Kobo$new(base_url_v2, base_url_v1, token)
kobo
```

For alternative ways of initializing the `Kobo` class, see [here](/articles/koboclient-class.html#the-use-of-koboclient-in-the-kobo-interface).

# Working with the Kobo class 

A good first step is to get an overview over all your assets. Almost everything in KoBoToolbox is an asset: questions, blocks, forms, templates. 

```{r include=FALSE}
vcr::use_cassette("get-assets", {
  assets <- kobo$get_assets()
})
```

```{r eval=FALSE}
assets <- kobo$get_assets()
```

```{r}
table(assets$asset_type)
```

You can also just get your forms/surveys: 

```{r include=FALSE}
vcr::use_cassette("get-surveys", {
  surveys <- kobo$get_surveys()
})
```

```{r eval=FALSE}
surveys <- kobo$get_surveys()
```

Inspect the data manually: 
```{r}
head(surveys)
```

You can then extract the ID of a single survey by filtering for its name. 

```{r}
test_id <- surveys %>% filter(name == "kbtbr Testing Survey") %>% pull(uid)
```

Get the answers (or _submissions_ in KoBoToolbox API lingo) to the form/survey:

```{r include=FALSE}
vcr::use_cassette("get-submissions", {
  submissions <- kobo$get_submissions(test_id)
})
```

```{r eval=FALSE}
submissions <- kobo$get_submissions(test_id)
```

```{r}
str(submissions)
```

Check out [Data wrangling of responses](/articles/wrangling-survey-responses.html) to see how to handle messy column names. 

# Kobo Administration with `kbtbr`

With `kbtbr`, you can also manage your assets and do other useful administration tasks right from your R console. 


## Create an Asset
You can also create assets using the `create_asset()` method of the Kobo class. This can be useful if you need to create a lot of assets at once.

```{r include=FALSE}
vcr::use_cassette("create-asset", {
  res <- kobo$create_asset("Question block", 
                  asset_type = "block", 
                  description = "this question block was created with the API.")
})

```

```{r eval=FALSE}
res <- kobo$create_asset("Question block", 
                  asset_type = "block", 
                  description = "this question block was created with the API.")
```

```{r}
res$status_http()
response_content <- res$parse("UTF-8") %>% jsonlite::fromJSON()
str(response_content, max.level = 1)
```


## Cloning an Asset

In order to clone an asset, you just have to specify its ID, a new name for the cloned asset and its type. 

```{r include=FALSE}
vcr::use_cassette("clone-asset", {
  res <- kobo$clone_asset(test_id, "Test survey copy", "survey")
})
```

```{r eval=FALSE}
res <- kobo$clone_asset(test_id, "Test survey copy", "survey")
```

```{r}
class(res)
res$status_http()
```

Parse out the ID of the asset that was created:

```{r}
response_content <- res$parse("UTF-8") %>% jsonlite::fromJSON()
str(response_content, max.level = 1)
test_clone_id <- response_content$uid
```


## Deploying an asset

To deploy an existing asset, e.g. after cloning or updating:

```{r include=FALSE}
vcr::use_cassette("deploy-asset", {
  res <- kobo$deploy_asset(test_clone_id)
})
```

```{r eval=FALSE}
res <- kobo$deploy_asset(test_clone_id)
```

```{r}
class(res)
res$status_http()
response_content <- res$parse("UTF-8") %>% jsonlite::fromJSON()
str(response_content, max.level = 1)
```

## Importing XLS forms

You can import an XLS form. This will create a draft survey/form which you could then again deploy using `kobo$deploy_asset()`. 

```{r include=FALSE}
vcr::use_cassette("import-xls-form", {
  res <- kobo$import_xls_form(name = "XLS Form via import", "assets/xls_form_downloaded.xls")
})
```
```{r eval=FALSE}
res <- kobo$import_xls_form(name = "XLS Form via import", "assets/xls_form_downloaded.xls")
```

```{r}
res$status_http()
```

# Making it your own: `post` and `get`

`kbtbr` is by no means complete. So you might find that some of your use cases are not covered by the existing methods of the `Kobo` class. For this purpose, `Kobo` as well as the lower-level `KoboClient` class expose generic `get` and `post` functions which should allow you to implement more functionalities on your own. You can check out the existing high-level functions (e.g. `get_assets` or `clone_asset`) to get ideas how your code should look like. Of course, contributions to `kbtbr` are always welcome. 


