---
title: "Advanced: The KoboClient class"
date: "`r Sys.Date()`"
output:  rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Advanced: The KoboClient class}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=TRUE, echo=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  echo = TRUE
)

# setup vcr
library(vcr)
vcr_dir <- "./assets"
if (!nzchar(Sys.getenv("KBTBR_TOKEN"))) {
    if (dir.exists(vcr_dir)) {
        # Fake API token to fool our package
        Sys.setenv("KBTBR_TOKEN" = "fakebearertoken")
    } else {
        # If there's no mock files nor API token, impossible to run tests
        stop("No API key nor cassettes, tests cannot be run.",
            call. = FALSE
        )
    }
}
invisible(vcr::vcr_configure(
    dir = vcr_dir,
    filter_request_headers = list(Authorization = "fakebearertoken")
))
vcr::check_cassette_names()
```

`KoboClient` is an R6 class that manages the connection between R and the server that hosts the KoboToolbox. 
Most users should not need the `KoboClient` class because `Kobo` takes care of the end-user services. `Kobo` uses `KoboClient` under the hood.

`KoboClient` is effectively an HTTP client and inherits its features from the 
[crul](https://github.com/ropensci/crul) library, in particular from `crul::HttpClient`.

# Getting started

## Initialization
To start a session, one needs to know the URL to the KoboToolbox server and the
API token, a unique key that allows interaction with the Kobo server. You can get the API key in your account settings.

One possibility is to store the token as an environment variable. 
The `KoboClient` will check if the variable `KBTBR_TOKEN` is set during the initialization. Otherwise, one can pass the token to the initialization function. 

The first step to start a session is to create an object from the R6 class by calling the `$new` method. 

```{r}
library(kbtbr)
base_url <- "https://kobo.correlaid.org"
my_session <- KoboClient$new(base_url, kobo_token = "faketoken")
class(my_session)
```


## The power of inheritance
Thanks to the inheritance mechanism, `my_session` can provide all the features of 
[`crul::HttpClient`](https://docs.ropensci.org/crul/reference/HttpClient.html)
that is the parent R6 class.

For example, 

```{r}
my_session$print()
```
will show you several properties set in the HTTP client, like the headers, the 
URL and other properties. The `$print()` method is not implemented in the KoboClient class, but it is automatically inherited from its parent class.

KoboClient is called a child of `crul::HttpClient`,
the parent,  and by the inheritance mechanism can have all of its features
(methods and fields) of the parent, but it can also provide more features or
specialize some of the parent methods. In our case, for example, some features
are added to work effectively with the KoboToolbox API: it stores the API token
as private variable, or some method are further specialized like the `$get` 
method described in the next section.


## `get` Method

The get method is based on the method  [`$get()`](https://docs.ropensci.org/crul/reference/HttpClient.html#method-get)
of the parent class. 
Hence, the `get` method will query the APIs and return an R6 object of class `crul::HttpResponse`. 

```{r include=FALSE, results='hide'}
vcr::use_cassette("get-assets", {
  assets <- my_session$get(path = "api/v2/assets/", query = list(format = "json"))
})
```

```{r eval=FALSE}
assets <- my_session$get(path = "api/v2/assets/")
```

```{r}
str(assets, max.level = 1)
```


# The use of KoboClient in the Kobo class

The `Kobo` class uses `KoboClient` for its HTTP requests. 
Kobo can create a session by initializing a `KoboClient` instance itself based on `base_url_v2`. This is probably the easiest way to initialize a `Kobo` instance.

```{r}
kobo <- Kobo$new(base_url)
```

Alternatively, the user can pass a session - a `KoboClient` instance associated with an API base URL - via the `session_v2` argument of `Kobo$new()`. The Kobo instance will then use this session for all the communication with the servers.

```{r}
# passing a KoboClient instance / a session to Kobo
my_session <- KoboClient$new(base_url)
kobo <- Kobo$new(session_v2 = my_session)
```

If the user specifies `base_url_v1` as an argument to `Kobo$new()` to make use of functionality based on KoBoToolbox API v1, the initializer of `Kobo` will initialize a second `KoboClient` instance to communicate with the v1 version of the API. The `Kobo` instance will then have two sessions, one for v2 and one for v1.

```{r}
# specifying session for API v2 directly, but letting Kobo initialize the one for API v1
kobo <- Kobo$new(session_v2 = my_session, base_url_v1 = "kc.correlaid.org")
kobo
```



```{r include=FALSE}
vcr::use_cassette("get-assets", {
  assets <- kobo$get_assets()
})
```

```{r eval=FALSE}
assets <- kobo$get_assets()
```
```{r}
colnames(assets)
```
