---
title: "R6 Classes"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

based on [Advanced R - Chapter 14](https://adv-r.hadley.nz/r6.html). If not otherwise specified, all quotes are from this chapter. 

```{r}
library(R6)
```

# Copy-on-modify vs. modify-in-place

> R6 objects have reference semantics which means that they are modified in-place, not copied-on-modify. 

typically, R uses a paradigm called "copy-on-modify", i.e. it makes a copy whenever you  modify an object: 

```{r}
x <- c(1, 2, 3)
y <- x

# x and y point to the same object in-memory
# cf https://adv-r.hadley.nz/names-values.html#binding-basics
addr_x <- lobstr::obj_addr(x)
addr_y <- lobstr::obj_addr(y)

addr_x == addr_y # TRUE

# now when we modify y, R makes a copy! 
y[[3]] <- 4
addr_y_after_modify <- lobstr::obj_addr(y)
addr_y_after_modify
addr_y == addr_y_after_modify # FALSE 
```

R6 classes constitute an exemption to this rule because they are **modified in place** and use **reference semantics**. This is not something we often run into when using R except in certain circumstances (see [here](https://adv-r.hadley.nz/names-values.html#modify-in-place)), so it might feel a bit foreign at first. 

```{r}
# this creates a class with a public field x which equals five 
myclass <- R6::R6Class(public = list(c = 5))

# we create two instances of our class 
instance_1 <- myclass$new() 
instance_2 <- instance_1

old_addr <- lobstr::obj_addr(instance_1) 
# instance_2 points to the same object in memory
lobstr::obj_addr(instance_1) == lobstr::obj_addr(instance_2) 

# modify the first instance
instance_1$c <- 2 * 5 # modify in place, not copy on modify!
instance_1$c # 10 
# instance_1 still stored at the old address -> R did not copy it somewhere else
lobstr::obj_addr(instance_1) == old_addr

# remember, we only changed instance_1 so far!!
instance_2$c # 10, huh? reference semantics!
old_addr == lobstr::obj_addr(instance_2)
```
Check out [the section on environments](https://adv-r.hadley.nz/names-values.html#env-modify) for a visual representation of what is happening here (replace `e1` and `e2` with `instance_1` and `instance_2` in your head).


# Classes, fields and methods


## Defining a class
> R6 only needs a single function call to create both the class and its methods: R6::R6Class(). This is the only function from the package that you’ll ever use!

### the first argument: `classname`

- not needed but improves error messages 
- in UpperCamelCase

```{r}
Client <- R6::R6Class(classname = "KoboClient")
Client
```

### second argument: public

contains the list of ...

- fields 
- methods (aka R functions)

in `snake_case`. It is implemented as a [named list](https://www.dummies.com/programming/r/how-to-create-a-list-in-r/). 

We can access the methods and fields using the `self$`. 
```{r public-argument}
ClientGenerator <- R6::R6Class(
  classname = "KoboClient",
  public = list(
    user_first_name = "Correl",
    user_last_name = "Aid",
    greeting_count = 0,
    greet = function(greeting = "Hi") {
      print(paste(greeting, self$user_first_name, self$user_last_name))
      self$greeting_count <- self$greeting_count + 1
      invisible(self)
    }
  )
)
```

> You should always assign the result of R6Class() into a variable with the same name as the class, because R6Class() returns an R6 object that defines the class

If we look at the Client object, it is a special "object generator" object that apparently can create `KoboClient` objects: `r emo::ji("eyes")`

```{r client-type}
ClientGenerator
```

We can now use the `new()` method of this object to make new **instances** of our class:

```{r instance-creation}
o1 <- ClientGenerator$new()
o1$greet("Hi")
```

## `$print()` and `$initialize()`

We can define a custom `$print()` method for our class: 

```{r}
ClientGenerator <- R6::R6Class(
  classname = "KoboClient",
  public = list(
    user_first_name = "Correl",
    user_last_name = "Aid",
    greeting_count = 0,
    greet = function(greeting = "Hi") {
      print(paste(greeting, self$user_first_name, self$user_last_name))
      self$greeting_count <- self$greeting_count + 1
      invisible(self)
    },
    print = function() {
      cat("Client: \n")
      cat("  User: ", self$user_first_name, "\n", sep = "")
    }
  )
)
o <- ClientGenerator$new()

```

with `$initialize()`, we can override the behavior of `$new()` so that users can also pass arguments when creating an instance:

```{r}
# this errors with the old class
# o <- ClientGenerator$new(user_first_name = "Correliiiii")
ClientGenerator <- R6::R6Class(
  classname = "KoboClient",
  public = list(
    user_first_name = NA,
    user_last_name = "Aid",
    greeting_count = 0,
    initialize = function(first_name = "Correl") { # we can set a default argument
      # we can also have validity checks here
      self$user_first_name <- first_name
    },
    print = function() {
      cat("Client: \n")
      cat("  User: ", self$user_first_name, "\n", sep = "")
      invisible(self)
    },
    greet = function(greeting = "Hi") { 
      print(paste(greeting, self$user_first_name, self$user_last_name))
      self$greeting_count <- self$greeting_count + 1
      invisible(self)
    }
  )
)
o <- ClientGenerator$new(first_name = "Correliiiii")
o$greet("Hello")
o$greeting_count
```

## Why do we return `invisible(self)`?

For methods that are mostly called for side-effects, we should always return the object itself invisibly. This allows for **method chaining** (very similar to the pipe!): 

```{r}
o <- ClientGenerator$new()
o$greet("Hello")$greet("Hi")$greet("Hey")$greeting_count

# or.. even more pipe like :) 
o$
  greet("Hello again")$
  greet("Hi again")$
  greet("Hey again")$
  greeting_count

```

