---
title: "Using crul with KoboToolbox"
output:
  github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval=FALSE
)
```

```{r setup, include=TRUE}
box::use(
  magrittr[`%>%`],
  dplyr[...],
  crul[...]
)
```

> **Note:** This article is intended to be excluded from a final state of
> of the project.


This WIP article develops a short overview on how to interact with the KoboToolbox
API with the plain [`{crul}`](https://github.com/ropensci/crul)
package, without any generalization etc. but purely to play around with
the provided functionality.

To replicate, please provide the following environment variables (e.g.
via your `~.Renviron` file):

- `KBTBR_BASE_URL`: something like `kobo.yourdomain.com`
- `KBTBR_TOKEN`: the token for your user. See 
  [here](https://support.kobotoolbox.org/api.html) on how to retrieve it.
  (currently not used in the examples)



## Basic requests

In the following, for demonstration purposes, we will show how to access the
API with `crul` on different ways. Those are not necessarily the one we chose
in the package, we just want to get a feeling for how everything works.

- Plain request: building a client object "by hand", executing its methods,
  parsing the results.
- Wrapping it in a function: avoid code repetition by hiding some of the
  previous steps in a general function.


We look at "low-hanging" endpoints: `assets` and `assets/{{uid}}/data`
(`v2` API only).


## Plain requests

```{r plain-crul}
base_url <- Sys.getenv("KBTBR_BASE_URL")
token <- Sys.getenv("KBTBR_TOKEN")

## Simple GET request

# Build the URL from a named-list style:
asset_url <- crul::url_build(
  url = base_url,
  path = "api/v2/assets",
  query = list(format = "json")
)

#' crul also has the reverse operation "url_parse" which decomposes an URL into
#' its elements this could be useful when "navigating" around the API using the
#' asset urls provided in the responses, e.g. in case we wanted to add elements
#' of it to a data frame representation etc.
crul::url_parse(asset_url)

# Get an instance of a crul::HttpClient:
crul_client <- crul::HttpClient$new(
  url = asset_url,
  headers = list(
    Authorization = paste0("Token ", token)
  )
)

crul_resp <- crul_client$get()
sloop::otype(crul_resp)
crul_resp$raise_for_status() # equivalent of httr::stop_for_status()

# Parse and inspect response
parsed_assets <- crul_resp$parse("UTF-8") %>%
  jsonlite::fromJSON()

# Inspect the results
names(parsed_assets)
str(parsed_assets, 1)
```


In the above example, we constructed the URL for the endpoint as follows:
- The base url (`kobo.example.com`) was queried via a secret token
- An additional path (`api/v2/assets`) and a query parameter (`format=json`)
  was passed as the final url was constructed using `crul::build_url()`.
- This final url was then passed as-is to the `crul::HttpClient` during
  instantiation.

However, it is possible (and more suitable/flexible)
to pass the additional path and query parameters during the actual method call
on the client, as is shown in the following example:


```{r base-url-client}
# Get an instance of a crul HttpClient:
crul_client <- crul::HttpClient$new(
  url = base_url,
  headers = list(
    Authorization = paste0("Token ", token),
    "content-type" = "application/json"
  )
)

# Make a GET request, and pass additional path and query parameters:
crul_resp2 <- crul_client$get(
  path = "api/v2/assets",
  query = list(format = "json")
  )

```


_Remark:_ Interestingly, it does not seem to matter whether which one we
use as the final passed url, both
- `https:://kobo.example.com/assets/?format=json` and
- `https:://kobo.example.com/api/v2/assets/?format=json`

This seems to be due to a legacy release of Kobo's
`v2` API, which probably used the shorter url form (which should not be used). 


### `POST` requests


In the first example, we can try to **create a new asset** (without any content)
via a `POST` request.

```{r}
# A little helper function to create a json-like string from
# a list.
# _note_ Kobotoolbox seems to be unable to deal with arrays, thus we have
# to unbox vectors of length 1.
list_as_json_char <- function(list) {
  jsonlite::toJSON(x = list, pretty = TRUE, auto_unbox = TRUE) %>%
    as.character()
}


body_create_example <- list_as_json_char(list(
  "name" = "A survey object created via API/R",
  "asset_type" = "survey"
))

# Exectute the request, and use method chaining to parse it directly 
crul_client$
  post(
    path = "api/v2/assets/",
    body = body_create_example)$
  parse("utf-8")

```

_Note_ somehow the additional nested parameter in "settings" resulted in a
failure of the request. It is questionable why they are in the documentation.
We cannot explain this _misbehaviour_ at the moment.


In the next example, in a similar style we **clone** an existing asset. This
needs two arguments passed in the body of the POST request.

```{r}
example_clone_body <- list_as_json_char(list(
  "clone_from" = "ajzghKK6NELaixPQqsm49e",
  "name" = "This is a cloned survey (via API/R)",
  "asset_type" = "survey"
))

crul_client$
  post(
    path = "api/v2/assets/",
    body = example_clone_body)$
  parse("utf-8")
```


### `PATCH` Requests


The newly created assets of type `survey` are initially in draft mode.
However, can we also move them to deployed?

Let's say we want to deploy all surveys which we have created beforehand,
and that we can identify because we put `API/R` in the name (of course,
this is nothing we would do in the package).

**DOES NOT WORK YET**

```{r}
# First, get again a list of assets
df_selected_uids <- crul_client$
  get(
    path = "api/v2/assets",
    query = list(formant = "json"))$
  parse("utf-8") %>%
  jsonlite::fromJSON() %>%
  purrr::pluck("results") %>%
  tibble::as_tibble() %>%
  dplyr::filter(stringr::str_detect(name, "API/R")) %>%
  dplyr::select(name, uid)

df_selected_uids

df_selected_uids$uid %>%
  purrr::walk(.f = function(uid) {
    crul_client$
      patch(
        path = glue::glue("api/v2/assets/{uid}/deployment/"),
        query = list(active = "false")
      )
  })

```



## Wrapping it in a function

Just to play around with possibilities, one can create a function that
consumes a `HttpClient` object and then performes a request on it, plus 
some additional data parsing / retrieval.

```{r wrap-in-function}
# Consider the following example client object
example_client <- crul::HttpClient$new(
  url = Sys.getenv("KBTBR_BASE_URL"),
  headers = list(Authorization = paste0("Token ", Sys.getenv("KBTBR_TOKEN")))
)


#' Simple request
#' 
#' Function that returns the parsed, tibbelized result from a simple GET
#' request
#' @param client An object of class HttpClient
#' @param path <string>
#' @param query <list> A named list of strings, giving additional request
#'  parameters.
#' @return A tibble
simple_request <- function(client, path, query = NULL) {

  stopifnot("HttpClient" %in% class(client))
  r <- client$get(path, query)
  r$raise_for_status()
  r$parse("UTF-8") %>%
    jsonlite::fromJSON() %>%
    purrr::pluck("results") %>%
    tibble::as_tibble()
}

simple_request(example_client, "api/v2/collections", list(format = "json"))
```

Consider a more complex example, where we first need to retrieve a list of
asset `uid` s, which we use in a second step to construct URLs to get
actually response data from survey assets.

```{r complexer-example}
assets_df <- simple_request(example_client, "api/v2/assets", list(format = "json"))

data_url <- assets_df %>%
  dplyr::filter(asset_type == "survey") %>%
  pull(data) %>%
  urltools::path()

full_dta_list <- data_url %>%
  purrr::map(function(url) {
    tryCatch(
      simple_request(example_client, path = url, query = list(format = "json")),
      error = function(e) {
        warning("Failed for {url}: {e}")
        return(NULL)
       })})

str(full_dta_list, 1)
```

_Note:_ you might have mentioned the somewhat awkward removing of the base
url and query, and then adding it implicit again. This is because the
client in this case always as a base url, and cannot tell (yet) whether it
was accidentially already prepended to the supplied `path`. We could
think about using `urltools::path()` somewhere to be sure to have everything
stripped away, given that the path does not start with `api/v2`... 
Has to be discussed.
